---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# saniabidjan

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)
[![R-CMD-check](https://github.com/openwashdata/saniabidjan/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/openwashdata/saniabidjan/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of saniabidjan is to provide survey data about the behavior and practices of managing sanitation facilties in two low-income areas of Abidjan, Cote d’Ivoire. 

## Installation

You can install the development version of saniabidjan from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/saniabidjan")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/openwashdata/saniabidjan/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to one dataset `saniabidjan`. It contains household survey data in two low-income areas of Abidjan. The focus of the survey is to understand the behavior and practices of the population in terms of managing sanitation facilities, as well as to assess the diseases they face.

```{r}
library(saniabidjan)
```

### saniabidjan

The dataset `saniabidjan` contains data about household surveys that cover 101 households in Abobo and 106 households in Yopougon. It has `r nrow(saniabidjan)` observations and `r ncol(saniabidjan)` variables

```{r}
saniabidjan |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "saniabidjan.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```


## Example

```{r}
library(saniabidjan)
library(ggplot2)
# Provide some example code here
saniabidjan |>
  ggplot(mapping = aes(x = facilities_type)) + 
  geom_bar(aes(fill = lat_care), position = "dodge") +
  labs(title = "Distribution of latrine care for different latrine types", 
       x = "Type of facilities",
       fill = "Latrine care type") 
```

## Capstone Project
This dataset is shared as part of a capstone project in Data Science for openwashdata. For more information about the project and to explore further insights, please visit the project page at https://ds4owd-001.github.io/project-coulbyph/

## License

Data are available as
[CC-BY](https://github.com/openwashdata/saniabidjan/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("saniabidjan")
```
